---
format:
  pdf: 
    output-file: "qmd_output.pdf"
    geometry: 
      - top=20mm
      - bottom=20mm
      - left=20mm
      - right=20mm
params:
  scenarios:  !expr c("ssp245", "ssp370")
  gridmet: !expr c("tmmx", "tmmn")
  cmip: !expr c("tas", "tasmax")
  location_id: "30001"
  location_name: "Test Location" 
---

# Climate Report for `r params$location_name`
```{r setup, echo=FALSE}
source("./plot.R")
source("./global.R")

get_historical <- function(id, element) {
  glue::glue(
    "{API_URL}/data/historical/{id}/{element}/"
  ) %>% 
    readr::read_csv(show_col_types = FALSE) 
}

get_future <- function(id, variable, type, scenarios){
  scenarios = paste(scenarios, collapse = ",")
  dat <- glue::glue(
    "{API_URL}/data/future/{id}/{variable}/"
  ) %>%
    httr::GET(
      query = list(
        table_type = type,
        scenarios = scenarios
      )
    ) %>% 
    httr::content(show_col_types = FALSE) %>%
    factor_scenario() 
}

calc_historical_pvalue <- function(timescale, dat) {
    if (timescale == "Annual") {
      dat <- dat %>% 
        dplyr::mutate(date = lubridate::floor_date(date, "year")) %>%
        dplyr::group_by(date) %>% 
        dplyr::summarise(
          value = dplyr::if_else(
            dplyr::first(variable %in% c("pr", "pet", "etr")), 
            sum(value), 
            mean(value)
          ) 
        ) 
    } else {
      dat <- dat %>% 
        dplyr::filter(which(month.name == timescale) == lubridate::month(date)) 
    }

   fit <- lm(dat$value ~ dat$date) %>% 
      summary() %>%
      purrr::pluck("coefficients") 
    
   return(tibble::tibble(
     p_value = purrr::pluck(fit, -1),
     slope = purrr::pluck(fit, 2)
   ) %>% 
     dplyr::mutate(
       p_value = p_value <=0.05,
       slope = slope * 10,
       timescale = timescale
     ))
}

historical_interpretive_text <- function(dat) {

  p_values <- c("Annual", month.name) %>% 
    magrittr::set_names(c("Annual", month.name)) %>% 
    purrr::map_df(calc_historical_pvalue, dat=dat)
  
  txt = "Between 1979 and 2020, there has been"
  if (!dplyr::filter(p_values, timescale == "Annual") %$% p_value) {
    txt = glue::glue("{txt} no statistically significant change in {tolower(legend_title(dat$variable[[1]], FALSE))} in {dat$name[[1]]} at an annual timescale.")
  } else {
    change <- dplyr::filter(p_values, timescale == "Annual") %$%
      round(slope, 4)
    txt = glue::glue("{txt} a statistically significant change in {tolower(legend_title(dat$variable[[1]], FALSE))} of {change} {text_units(dat$variable[[1]])} per decade at an annual timescale.")
  }
  
  filt <- p_values %>% 
    dplyr::filter(p_value) %>%
    dplyr::rowwise() %>% 
    dplyr::mutate(text = glue::glue("{timescale} ({round(slope,4)} {text_units(dat$variable[[1]])} per decade)")) %>% 
    dplyr::filter(timescale != "Annual")
  
  if (nrow(filt) == 0) {
    txt <- glue::glue("{txt} When looking at monthly data, there has also been no significant trend in {tolower(legend_title(dat$variable[[1]], FALSE))} across the same time period.")
  } else {
    txt <- glue::glue("{txt} When looking at monthly data, there have been significant changes in {tolower(legend_title(dat$variable[[1]], FALSE))} in {knitr::combine_words(filt$text)}.")
  }
  
  txt <- glue::glue("{txt} Below, annual trends in {tolower(legend_title(dat$variable[[1]], FALSE))} are plotted for the GridMet period of record.")
  return(txt)
}

future_monthly_table <- function(dat) {
    avg <-  dat %>% 
      dplyr::select(-upper, -lower, -grp) %>% 
      dplyr::distinct() %>% 
      dplyr::filter(scenario == "Historical Emissions") %>% 
      dplyr::select(month, avg=value)
    
    dat %<>%
      dplyr::filter(scenario != "Historical Emissions") %>% 
      dplyr::left_join(avg, by="month") %>% 
      dplyr::mutate(
        upper = upper - avg, 
        lower = lower - avg, 
        value = value - avg
      )
    
    dat %>% 
      dplyr::select(
        Scenario = scenario,
        Month = month,
        Difference = value, 
        `Time Period` = grp
      ) %>% 
      dplyr::mutate(
        Month = factor(Month, levels = month.abb)
      ) %>% 
      tidyr::pivot_wider(names_from = Scenario, values_from = Difference) %>% 
      dplyr::arrange(dplyr::desc(`Time Period`), Month) %>% 
      dplyr::select(-`Time Period`) %>% 
      knitr::kable(booktabs=TRUE) %>% 
      kableExtra::pack_rows(
        index = c(
          "Mid Century (2040-2069)" = 12,
          "End-of-Century (2070-2099)" = 12
        )
      )
}

future_monthly_text <- function(dat) {
    v = dat$variable[[1]]
    avg <-  dat %>% 
      dplyr::select(-upper, -lower, -grp) %>% 
      dplyr::distinct() %>% 
      dplyr::filter(scenario == "Historical Emissions") %>% 
      dplyr::select(month, avg=value)
    
    dat %<>%
      dplyr::filter(scenario != "Historical Emissions") %>% 
      dplyr::left_join(avg, by="month") %>% 
      dplyr::mutate(
        upper = upper - avg, 
        lower = lower - avg, 
        value = value - avg
      ) %>% 
      dplyr::group_by(scenario, grp) %>% 
      dplyr::slice(which.max(abs(value)))%>%
      dplyr::mutate(
        scenario = stringr::str_split(scenario, "\n") %>% 
                      unlist() %>% 
                      purrr::pluck(1)
      ) %>%
      
      dat%>% dplyr::mutate(txt = glue::glue("{month} is projected to see the largest {ifelse(value > 0, 'increase', 'decrease')} in {tolower(legend_title(v, FALSE))} ({round(value, 2)} {text_units(v)}) under the {tolower(scenario)} scenario relative to the 1991 - 2020 baseline.")) %>%
      dplyr::group_by(grp) %>% 
      dplyr::summarise(txt = knitr::combine_words(txt)) %>% 
      dplyr::mutate(txt = glue::glue("By {tolower(grp)}, {txt}"))
    
    txt = glue::glue("In addition to changing at annual time scales, {tolower(legend_title(v, FALSE))} is also projected to change at the monthly scale. {dplyr::filter(dat, grp == 'Mid Century (2040-2069)') %$% txt} {{dplyr::filter(dat, grp == 'End-of-Century (2070-2099)') %$% txt}}. Below, a table and graph showing changes in {tolower(legend_title(v, FALSE))} for {lapply(params$scenarios, scenario_switch) %>% knitr::combine_words()} scenarios is provided to give a monthly breakdown of projected changes.")
}

future_timeseries_text <- function(dat) {

  groups = c("Middle of the Road\n(SSP2-4.5)",
            "High Emissions\n(SSP3-7.0)")
  
  p_vals <- groups %>% 
    purrr::map_df(function(x) {
      tmp <- dplyr::filter(
        dat,
        scenario %in% c(x, "Historical Emissions")
      ) %>% 
        dplyr::arrange(year)
      
     fit <- lm(tmp$value ~ tmp$year) %>% 
        summary() %>%
        purrr::pluck("coefficients") 
      
     return(tibble::tibble(
       p_value = purrr::pluck(fit, -1),
       slope = purrr::pluck(fit, 2)
     ) %>% 
       dplyr::mutate(
         p_value = p_value <=0.05,
         slope = slope * 10,
         scenario = x
       ))
  }) %>% 
    dplyr::rowwise() %>% 
    dplyr::filter(p_value) %>% 
    dplyr::rowwise() %>%
    dplyr::mutate(
      scenario = stringr::str_split(scenario, "\n") %>% 
                    unlist() %>% 
                    purrr::pluck(1),
      txt = glue::glue("under the {tolower(scenario)} emissions scenario, {tolower(legend_title(dat$variable[[1]]))} will {ifelse(slope > 0, 'increase', 'decrease')} at a rate of {round(slope, 2)} {text_units(dat$variable[[1]])} per decade")
    )
  
  txt = "Between 1950 and 2099, it is projected that"
  if (length(p_vals) == 0) {
    txt = glue::glue("{txt} there will be no statistically significant changes in {tolower(legend_title(dat$variable[[1]]))} in neither {lapply(params$scenarios, scenario_switch) %>% knitr::combine_words(and = ' nor ')}")
  } else {
    txt = glue::glue("{txt} {knitr::combine_words(p_vals$txt)}.")
  }
  
  txt = glue::glue("Below, a timeseries plot shows projected changes in  {tolower(legend_title(dat$variable[[1]]))} under these different emission scenarios.")
```
 
 
```{r}
  p_values <- c("Annual", month.name) %>% 
    magrittr::set_names(c("Annual", month.name)) %>% 
    purrr::map_df(calc_historical_pvalue, dat=dat)
  
  txt = "Between 1979 and 2020, there has been"
  if (!dplyr::filter(p_values, timescale == "Annual") %$% p_value) {
    txt = glue::glue("{txt} no statistically significant change in {tolower(legend_title(dat$variable[[1]], FALSE))} in {dat$name[[1]]} at an annual timescale.")
  } else {
    change <- dplyr::filter(p_values, timescale == "Annual") %$%
      round(slope, 4)
    txt = glue::glue("{txt} a statistically significant change in {tolower(legend_title(dat$variable[[1]], FALSE))} of {change} {text_units(dat$variable[[1]])} per decade at an annual timescale.")
  }
  
  filt <- p_values %>% 
    dplyr::filter(p_value) %>%
    dplyr::rowwise() %>% 
    dplyr::mutate(text = glue::glue("{timescale} ({round(slope,4)} {text_units(dat$variable[[1]])} per decade)")) %>% 
    dplyr::filter(timescale != "Annual")
  
  if (nrow(filt) == 0) {
    txt <- glue::glue("{txt} When looking at monthly data, there has also been no significant trend in {tolower(legend_title(dat$variable[[1]], FALSE))} across the same time period.")
  } else {
    txt <- glue::glue("{txt} When looking at monthly data, there have been significant changes in {tolower(legend_title(dat$variable[[1]], FALSE))} in {knitr::combine_words(filt$text)}.")
  }
  
  txt <- glue::glue("{txt} Below, annual trends in {tolower(legend_title(dat$variable[[1]], FALSE))} are plotted for the GridMet period of record.")
  return(txt)
}

plot_future <- function(dat, type) {
  if (type == "monthly") {
    plt <- dat %>% 
      dplyr::mutate(month = factor(month, levels = month.abb)) %>%
      make_monthly_plot(TRUE, FALSE) 
  } else {
    plt <- make_timeseries_plot(dat, TRUE, FALSE)
  }
  return(plt)
}

historical_df <- tibble::tibble(
  variable = params$gridmet
) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(
    df = list(get_historical(params$location_id, variable)),
    plots = list(make_historical_plot(df, variable, "Annual"))
  )


future_df <- tibble::tibble(
  variable = params$cmip
) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(
    monthly = list(get_future(params$location_id, variable, "monthly", params$scenarios)),
    timeseries = list(get_future(params$location_id, variable, "timeseries", params$scenarios))
  ) %>% 
  tidyr::pivot_longer(
    c("monthly", "timeseries"),
    names_to = "type",
    values_to = "df"
  ) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(
    plot = list(plot_future(df, type))
  )
```
# Historical Trends
Climate has changed over the last 50 years in `r params$location_name`. In this analysis, we use [GridMet](https://www.climatologylab.org/gridmet.html) meteorology to look at trends in climate in `r params$location_name` since 1979. We do this analysis for `r lapply(params$gridmet, function(x) {tolower(legend_title(x, FALSE))}) %>% knitr::combine_words()` in the section below.

``` {r}
#| output: asis
#| echo: false
res <- purrr::pmap_chr(historical_df, \(variable, df, plots) { 
  knitr::knit_child(text = c(
    "## Historical Trends in `r variable`",
    "",
    "`r historical_interpretive_text(df)`",
    "",
    "```{r}",
    "#| echo: false",
    "plots",
    "```",
    ""), envir = environment(), quiet = TRUE)
})
cat(res, sep = "\n")
```


# Climate Projections
It is projected that climate in `r params$location_name` will change considerably by the end of the century, depending on the emissions scenario.

## Timeseries of Projections
Here is a timeseries of how climate is projected to change by 2100:


## Monthly Climate Projections
There can be a large variability of how climate changes across the year. Below are plots of projected change in monthly conditions over the next century:

